Radioastronomical image reconstruction using Kalman filter under ionospheric perturbations

### 2022-08-24 article/article.tex
Update article.tex
almost there

\newcommand{\x}{\boldsymbol{x}}
\newcommand{\y}{\boldsymbol{y}}
\newcommand{\z}{\boldsymbol{z}}

@@

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.7\linewidth]{src/sites_geoo_satie}
		\caption{Geographic sites of SATIE (copyright E. Vourc'h)}
		\label{fig:sitesgeoosatie}
	\end{figure}

@@

	\begin{tabular}{|c||c|c|} % a traduire ffs
		\hline
		Divisions & Components and Systems for electrical energy & Multi-scale information and analysis systems \\ & (CSEE) & (SIAME) \\\hline
		Research fields & \begin{tabular}{@{}c@{}}
			Materials: from synthesis to application\vspace{8px} \\ Power electronics: integration, \\ usage constraints, interfaces\vspace{8px} \\ Transducers and energy systems: \\ methodologies, design and control
		\end{tabular} & \begin{tabular}{@{}c@{}}
			Instrumentation and imagery\vspace{5px} \\ Methods and tools for signals and systems\\ (MOSS)
		\end{tabular}
		\\\hline
	\end{tabular}
	\vspace{7px}\\
	I did my internship in the SIAME/MOSS division. \\
	The SIAME/MOSS division is focused on getting a better understanding and mastering complex systems. The three main topics of the division are :
	\paragraph{Data and image analysis:} this topic's main goal is to estimate physical or semantic quantities from observations data. This data is incomplete, inaccurate, and not absolutely reliable. When data is generated using multi-sensor systems, it presents similarities and complementarities that can be used to get better estimations.

@@

@@

	A (non extended) Kalman filter is a linear quadratic estimator, using observations, state-transition model and knowledge about noise and errors to compute a more accurate result than observations alone. Radio astronomical measures being prone to noise and errors, usage of this filtering method seems \emph{a priori} a good way to improve accuracy of reconstruction. \\

@@

@@

@@

@@

	Those inputs are not \emph{a priori} known, but it has been shown that using the \emph{autocovariance least-squares} (ALS) method provides a good estimate of those matrices. [CITATION NEEDED] \\
	We will assume that these inputs are known perfectly except when specified otherwise.\\

@@

	\paragraph{Error covariance matrices} : as mentioned above, these matrices set can be estimated with good accuracy using ALS algorithms [CITATION NEEDED]. It is to be noted that errors on these matrices will not induce new errors in the estimated images, as they are only used to compute the Kalman gain. This implies that inaccurate error matrices can only spread already existing errors on estimated images.

@@

@@

	The first interesting observations on these graphs are the extremities convergences : above 0,05m position error standard deviation, errors approach a distance of 0,1. The $d_1$ distance being normalised (as $\x$ and $\xe$ are normalised between 0 and 1), this distance means 10\% reconstruction error. This error level represents non reliable reconstruction. It can be illustrated by the following reconstruction :
	\begin{figure}[H]
		\centering
		\begin{subfigure}{.5\textwidth}
			\centering
			\includegraphics[width=.7\linewidth]{src/t_im.png}
			\caption{Errorless reconstruction}
		\end{subfigure}%
		\begin{subfigure}{.5\textwidth}
			\centering
			\includegraphics[width=.7\linewidth]{src/t_err.png}
			\caption{Reconstruction at 0,1 distance of the true image}
		\end{subfigure}
		\caption{Image reconstruction at unreliable distance}
	\end{figure}

@@

	Ionospheric perturbations being local and direction-dependent, it can be interesting to measure the effect of a single direction being inaccurate. To evaluate the reconstruction error, the $d_2$ distance is used, as it is additive: the total distance between two reconstructions is the sum of the distance of each direction.

@@

	\newpage
	\part{Experimental correction}
	\section{implementation}
	\section{Measures}

	\newpage
	\part{Conclusion}

	\newpage
	\part{Further insights}

@@

>>>>

\newcommand{\x}{\boldsymbol{x}}
\newcommand{\y}{\boldsymbol{y}}
\newcommand{\z}{\boldsymbol{z}}
\renewcommand{\u}{\boldsymbol{u}}

@@

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.7\linewidth]{src/sites_geoo_satie}
		\caption{Geographic sites of SATIE (\textcopyright E. Vourc'h)}
		\label{fig:sitesgeoosatie}
	\end{figure}

@@

	\begin{tabular}{|c||c|c|} % a traduire ffs
		\hline
		Divisions & Components and Systems for electrical energy & Multi-scale information and analysis systems \\ & (CSEE) & (SIAME) \\\hline
		Research fields & \begin{tabular}{@{}c@{}}
			Materials: from synthesis to application\vspace{8px} \\ Power electronics: integration, \\ usage constraints, interfaces\vspace{8px} \\ Transducers and energy systems: \\ methodologies, design and control
		\end{tabular} & \begin{tabular}{@{}c@{}}
			Instrumentation and imagery\vspace{5px} \\ Methods and tools for signals and systems\\ (MOSS)
		\end{tabular}
		\\\hline
	\end{tabular}
	\vspace{7px}\\
	I did my internship in the SIAME/MOSS division. \\
	The SIAME/MOSS division is focused on getting a better understanding and mastering complex systems. The three main topics of the division are :
	\paragraph{Data and image analysis:} this topic's main goal is to estimate physical or semantic quantities from observations data. This data is incomplete, inaccurate, and not absolutely reliable. When data is generated using multi-sensor systems, it presents similarities and complementarities that can be used to get better estimations. \\
	The applications are :
	\begin{itemize}
		\item Geolocation
		\item Non-destructive testing
		\item radio telescopes calibration
		\item image analysis for shape detection
	\end{itemize}

	\paragraph{Algorithms-architectures-uses adequacy:} This topic's goal is the design of autonomous systems with matching algorithms and algorithms. They develop tools and methods for multi-sensor instrumentation, embedded computing architectures and data flux synchronisation. Their work is divided in 4 points :

	\paragraph{Perception systems for mobility:} The last research axis is mostly aimed at the development of autonomous vehicles. Its goal is to conceive the hardware part of a perception system, and the software analysis sequence to optimise the system. \\

	The SATIE places a great importance on experimentation, and notably concerning the autonomous vehicles conception, as all results must be confirmed with experimentation, the lab has many electrical cars mounted with instruments and automated controls. \\

	This organisation can be summarised as follows :

	\begin{figure}[H]
		\centering
		\includegraphics[width=.7\linewidth]{src/satie_org}
		\caption{lab's organisation}
	\end{figure}

	The lab is also involved in many masters in Université Paris-Saclay, ENS Paris-Saclay, or ENS Rennes etc. These masters cover the topics: signal processing; electronics, sensors; electrical energy; mechatronics and higher education (M2FESUP). It also accepts some PhD students and interns.

@@

	\begin{figure}[H]
		\centering
		\includegraphics[width=.5\linewidth]{src/radioastro}
		\caption{Radio telescope in New Mexico, Very Large Array (VLG)}
	\end{figure}

@@

	A Kalman filter is a linear quadratic estimator, using observations, state-transition model and knowledge about noise and errors to compute a more accurate result than observations alone. Radio astronomical measures being prone to noise and errors, usage of this filtering method seems \emph{a priori} a good way to improve accuracy of reconstruction. \\

@@

	As of today, many optimisations have been done to improve Kalman filter, such as smoothers \cite{RTS}, and models to describe accretion disks evolution. This raises the question about the feasibility of using a Kalman filter to enhance radio astronomic pictures of dynamic systems.

@@

	Similar to $d_1$, but better known, to measure the quality of image reconstruction, the Peak to Signal Noise Ratio (PSNR) can be computed :

	$$
		PSNR = 10\cdot\log_{10}\left(\frac{d^2}{EQM}\right)
	$$

	where $EQM$ is the mean quadratic error between reference image $I_{ref}$ and reconstruction $I_{rec}$ :

	$$
		EQM = \frac{1}{M_xM_y}\sum_{i=0}^{M_x-1}\sum_{j=0}^{M_y-1}( I_{ref}(i,j) - I_{rec}(i,j) )^2
	$$

	$d$ is the dynamic of each pixel (here 1), $M_X$ and $M_Y$ the size of the image along respectively axis $X$ and $Y$ in pixels.

@@

	\begin{figure}[H]
		\centering
		\includegraphics[width=.5\linewidth]{src/antennes}
		\caption{Illustration of the antennas mesh}
	\end{figure}

@@

	Those inputs are not \emph{a priori} known, but it has been shown that using the \emph{autocovariance least-squares} (ALS) method provides a good estimate of those matrices. \cite{ALS} \\
	We will assume that these inputs are known perfectly except when specified otherwise.\\

@@

	\paragraph{Error covariance matrices} : as mentioned above, these matrices set can be estimated with good accuracy using ALS algorithms \cite{ALS}. It is to be noted that errors on these matrices will not induce new errors in the estimated images, as they are only used to compute the Kalman gain. This implies that inaccurate error matrices can only spread already existing errors on estimated images.

@@

	\subsection{Initialisation error}

	When using the Kalman filter, having the most accurate initialisation is important. This part has two goals: measure how an initialisation error affects reconstruction; and verify that the filter works correctly by eliminating these errors after a few snapshots. \\
	To measure this, every parameter is considered perfect except the initial $\x_0$ and $\P_0$ computed accordingly. This test was done using a $3\times 3$ image of a \mintinline{matlab}|linspace|\footnote{This was done before the methodology establishment}, to which a $90\deg$ rotation was applied each snapshot. The error on the initial image was introduced by changing the Signal to Noise Ratio (SNR) during reconstruction. \\

	The following reconstruction can be observed :

	\begin{figure}[H]
		\centering
		\includegraphics[width=.8\linewidth]{src/K_ev}
		\caption{Kalman filtering at different snapshots}
		\label{fig:K_ev}
	\end{figure}

	On this figure (\ref{fig:K_ev}), the top line represents the original images set with the first replaced with inaccurate reconstruction, and the bottom line is the Kalman reconstruction at the same snapshots, titles of bottom images are the PSNR of these measures. \\

	This figure illustrates how the initialisation error spreads between snapshots and how the filtering rapidly compensate it. This evolution can be plotted :

	\begin{figure}[H]
		\centering
		\includegraphics[width=.5\linewidth]{src/evPSNR}
		\caption{PSNR evolution of 1 simulation with inaccurate initialisation}
	\end{figure}

	On this figure, it can be observed that the error is compensated after 4 iterations. \\

	To determine the number of snapshots in which the initialisation error was not the most corrected, a set of measures was done using different SNR :

	\begin{figure}[H]
		\centering
		\includesvg[width=.65\linewidth]{graphs/SNR.svg}
		\caption{Number of iteration needed to reach the best PSNR}
	\end{figure}

	An important observation made during experimentation is that the number of snapshots needed was never observed over 5. It mostly stays around 4 and sometimes is 5, so this is the number of snapshots we will consider for stabilisation of initial error. \\
	The measure of the maximum PSNR can also be measured, but was not in this work.

@@

	The first interesting observations on these graphs are the extremities convergences : above 0,05m position error standard deviation, errors approach a distance of 0,1. The $d_1$ distance being normalised (as $\x$ and $\xe$ are normalised between 0 and 1), this distance means 10\% reconstruction error. This error level represents non reliable reconstruction. It can be illustrated by the following reconstruction :
	\begin{figure}[H]
		\centering
		\begin{subfigure}{.5\textwidth}
			\centering
			\includegraphics[width=.7\linewidth]{src/t_im.png}
			\caption{Errorless reconstruction}
		\end{subfigure}%
		\begin{subfigure}{.5\textwidth}
			\centering
			\includegraphics[width=.7\linewidth]{src/t_err.png}
			\caption{Reconstruction at 0,1 distance of the true image}
		\end{subfigure}
		\caption{Image reconstruction at unreliable distance}
		\label{fig:high_err}
	\end{figure}

@@

	Ionospheric perturbations being local and direction-dependent, it can be interesting to measure the effect of a single direction being inaccurate. To evaluate the reconstruction error, the $d_2$ distance is used, as it is additive: the total distance between two reconstructions is the sum of the distance of each direction. The maximum distance is thus not normalised and errors superior to 1 appear, these errors cannot be compared to directional errors above. The results are the following :

	\begin{figure}[H]
		\centering
		\begin{subfigure}{.5\textwidth}
			\centering
			\includesvg[width=.9\linewidth]{graphs/dir4_tot.svg}
			\caption{Distance to true image}
		\end{subfigure}%
		\begin{subfigure}{.5\textwidth}
			\centering
			\includesvg[width=.9\linewidth]{graphs/dir4_1.svg}
			\caption{Distance of inaccurate direction to the true direction}
		\end{subfigure}
		\caption{Simulation results for directional inaccuracy on one direction}
	\end{figure}

	The first notable observation is that the standard deviation of reconstruction error under directional perturbation is important, even after averaging 100 simulations : the reconstruction is thus more sensitive to directional perturbation. Another observation is that past the 0.05 direction error, the average reconstruction error is basically the same, with very small variations of the confidence interval. This is in fact the same case as figure (\ref{fig:high_err}), where the error is only the mean distance between two random images as the reconstruction is not reliable any more, being mostly errors. \\

	Another remarkable observation is that the error on the single direction seems to grow slightly faster than the total error. This can be verified when computing the average percentage of the single direction error in the total one :

	\begin{figure}[H]
		\centering
		\includesvg[width=.65\linewidth]{graphs/dir4_perc.svg}
		\caption{Part of the single direction reconstruction error in the total one}
	\end{figure}

	This figure also shows that the reconstruction error in the erroneous direction is more important than on the others as it was expected (out of 100 directions, the mean error should be 1\% in each direction). This indicates that prior knowledge on the error may help identifying outliers in order to ignore them if their value makes no sense in regard of the other directions.

@@

	To better understand the reconstruction error sensitivity to direction inaccuracies, the same simulation was done with errors on all directions, using $d_1$ distance for normalisation and reference to position error. The results are the following :

	\begin{figure}[H]
		\centering
		\begin{subfigure}{.5\textwidth}
			\centering
			\includesvg[width=.9\linewidth]{graphs/dirA_x_e.svg}
			\caption{Distance to true image}
		\end{subfigure}%
		\begin{subfigure}{.5\textwidth}
			\centering
			\includesvg[width=.9\linewidth]{graphs/dirA_xx.svg}
			\caption{Distance to unperturbed reconstruction}
		\end{subfigure}
		\caption{Simulation results for directional inaccuracy on all direction}
	\end{figure}

	It can be observed that the error tends faster towards the random error case seen in (\ref{fig:high_err}). This can be qualified as directional resolution was equal to $2/10=0.2$ (direction cosines) between two directions, and positional resolution was $1.472$m between two  antennas. With this taken into account, the reconstruction is still more sensitive to directional error than positional error.

	\newpage
	\part{Experimental correction}

	The filter being sensitive to directional error, the goal of this part is to use prior knowledge to estimate this error. The idea is to use a known $\y$, measured with the inaccessible $\H$, and a MVDR reconstruction $\x_{MVDR}$ \cite{TER} from this measure and solve the problem :

	% IDEA : tourner sans bouger x_MVDR
	% => ça marche pas

	$$
		\y - \H^d*\x_{MVDR} = 0
	$$
	Where :

	$$
		\H_{j,q}^d = \exp(-i\frac{2\pi}{\lambda}\Delta\z_j\dotproduct(\I_q^* - \u_q))
	$$

	The resulting vector $\u$ should be the correction vector such as :
	$$
		\I = \I^*-\u
	$$

	This problem being not linear, its analytic resolution is complicated. To solve it, it was implemented under Matlab using \mintinline{matlab}|lsqnonlin| function. This function does not support complex values, so the problem had to be broken up into real and complex parts (solved simultaneously), see code details in appendix.\\

	However, after numerous tests, solving this problem does not seem feasible, at least using this function and these parameters: all the tests presented a solution worse than simply keeping the inaccurate matrix. The average deterioration has been measured :

	$$
		\left\langle d_1(\I,\I^*) - d_1(\I,\I^*-\u)\right\rangle = -4.12\times 10^{-4}
	$$

	To alleviate this issue, introducing an \emph{a priori} on the initial solution has been tested, without notable amelioration: the deterioration was only slightly weaker. \\

	Some other solutions that were tried without being able to solve the problem:
	\begin{itemize}
		\item lowering function tolerance
		\item lowering/raising step size and tolerance
		\item switch from trust region reflective algorithm to Levenberg-Maquardt
		\item raising max iterations and max evaluations to allow the algorithm to reach solutions with these parameters
	\end{itemize}

	\newpage

	\part{Conclusion}

	The results of this study demonstrate multiple points regarding the usability of a Kalman filter in the radio astronomic context. The first conclusion is that errors in the initial picture, even in good conditions, can continue to spread errors up to the fifth snapshot. Then, antennas positions should be known with at least 0.5\% relative precision, as an inaccuracy superior to that will create non-negligible errors in the final reconstruction. Concerning directions, those must be known with at least 0.1\% relative precision for the same reason. \\

	Trying to estimate the error on directions could be a good way to get better results, however, analytic solution or a different resolution algorithm should be found, as both algorithms (trust region reflective and Levenberg-Maquardt) used by \mintinline{matlab}|lsqnonlin| does not produce conclusive results.

	\newpage
	\part{Recommended future works}

@@

	\newpage
	\part{Appendices}
	\section{Experimental correction code}

\begin{minted}{matlab}
% init
x0 = zeros(1,2*D);
lb = [-0.1*ones(1,2*D)];
ub = [0.1*ones(1,2*D)];

% problem definition
Iv = reshape(I_err,[],1); % I_err is I with errors
prb = @(x)prb_f(x,y,Iv,dz,J,D,lambda,X_mvdr); % creation of function f(x) = y-H(x)*x_MVDR with static inputs

% lsqnonlin
options = optimoptions(@lsqnonlin,'Display','Iter','FiniteDifferenceType','central'); % FiniteDifferenceType central computes on both sides of the gradient, did not help but a bit more precise
x = lsqnonlin(prb,x0,lb,ub,options); % solve f(x)=0

% result extraction
Icorr = - reshape(x(1:2*D),D,2) + I_err; % I_corr is I corrected
Hcorr = matF(J,D,z_err,lambda,Icorr); % matF computes H

.
.
.

% function definition at the end of the file
function s = prb_f(x,y,Iv,dz,J,D,lambda,xMVDR)
Hmodif = zeros(J^2,D);
for q=1:D
Hmodif(: ,q) = exp(-1i*2*pi/lambda*(dz(:,1)*(Iv(q) - x(q))+dz(:,2)*(Iv(D+q) - x(D+q))));
end 
ymodif = Hmodif*xMVDR;
s = [real(y);imag(y)] - [real(ymodif);imag(ymodif)];
end

\end{minted}
